Learning by Teaching a Robot:\\The Case of Handwriting

### 2016-01-04 rebuttal.tex

\section*{Reporting of the studies}

We have clarified our wording to only draw conclusion from factual measured
data or feedback from the parents/teachers/therapists (this is in particular
the case for the case-study 1),

We have added new quantitative data in the reporting of the case study 2, that
show a quantifiable improvement of handwriting. We want however to remain
cautious in our claims: handwriting acquisition is a complex cognitive skill.
While the child demonstrated clear improvements over the course of our
experiment, measuring the lasting effects of the robot-based remediation is
difficult (as many other individual and social factors may have impacted the
child's performance as well).

Finally, we have added the results of one more experiment (conducted during the
summer 2015), that involved 8 children with handwriting deficits. This
experiment focused on measuring how \emph{serious} the children were at
teaching the robot.

@@

\section*{Other comments}

One of the reviewers asked for clarification on a few other points:

\begin{itemize}

\item \emph{Did we include children who had difficulty with the dynamics of
turn-taking? If so, how did the system fare with those children?} In the case
the children did not pro-actively engage into turn-taking, the facilitator (be
it one of the experimenters or the therapist) would suggest it.

@@

\item \emph{The handwriting/letter generation is sufficiently described, but
how are the corresponding robot gestures generated?} We relied on NaoQI inverse
kinematics solver. This has been added to the text.

\item \emph{Was an interaction in which the child moves the hand of the NAO to
write the letter considered? Why did the authors choose the current interaction
model in which the child demonstrates on a tablet?} The Nao robot does not
feature force/torque sensing, which prevents most of the interesting direct
physical interaction like moving the hand of the robot. With appropriate
hardware, this interaction modality would be interesting to explore, though.

\item \emph{Could the measure of engagement plotted in Figure 7 be confounded
with the physically-active aspect of the hand-writing activity versus the
passive storytelling activity?} We have decided to replace this figure with an
objective metric (with-me-ness) that relies on the focus of attention of the
children to assess to which extent the child is into the task.

### 2016-01-04 rebuttal.tex

\begin{document}

Dear Editor, Dear Reviewers,

\vspace{2em}

We would like to thank both the reviewers and the editor for going through our
manuscript. As a follow-up of your comments, the manuscript underwent
several significant changes, that we summarize hereafter.

@@

While the reviews are generally positive (quoting one of the reviewer, ``this
is an excellent model for this kind of research, and the results are
promising''), we certainly take the two main concerns raised from the reviews:

\begin{itemize}
    \item The reporting of the studies is lacking
    \item Possible ethical concerns are not sufficiently discussed
\end{itemize}

@@

This revision of our manuscript attempts to address both these concerns.

\section*{Reporting of the studies}

We have clarified our wording to only draw conclusion from factual measured
data or feedback from the parents/teachers/therapists (this is in particular
the case for the case-study 1),

We have added quantitative data in the reporting of the case study 2, that
explicitly show a quantifiable improvement of handwriting. We want however to remain
cautious in our claims: handwriting acquisition is a complex cognitive skill.
While the child demonstrated clear improvements over the course of our
experiment, measuring the lasting effects of the robot-based remediation is
difficult (as many other individual and social factors may have impacted the
child's performance as well).

We have also added the results of an additional experiment (conducted during the
summer 2015), that involved 8 children with handwriting deficits. This
experiment focused on measuring how \emph{serious} the children were at
teaching the robot.

Finally, we have added a whole new section (section 4) that explicitly
summarizes and discusses our findings.

@@

\section*{Ethical concerns}

An explicit paragraph has been added to the conclusion that states the possible
ethical concerns raised by this work. Note that, while the psychological impact of the
mentor-protégé relationship has been studied in the literature, it has been
discussed from the perspective of the protégé only. In our situation, the protégé
is the robot, and the child is the mentor: to our best knowledge, the ethical
implications for the child of such a situation are mostly an open question. We
underline accordingly this point in the article.

@@

\item \emph{The handwriting/letter generation is sufficiently described, but
how are the corresponding robot gestures generated?} We relied on NaoQI inverse
kinematics solver. This has been added to the text.

\item \emph{Was an interaction in which the child moves the hand of the NAO to
write the letter considered? Why did the authors choose the current interaction
model in which the child demonstrates on a tablet?} The Nao robot does not
feature force/torque sensing, which prevents most of the interesting direct
physical interaction like moving the hand of the robot. With appropriate
hardware, this interaction modality would be interesting to explore, though.

\item \emph{Could the measure of engagement plotted in Figure 7 be confounded
with the physically-active aspect of the hand-writing activity versus the
passive storytelling activity?} We have decided to replace this figure with an
objective metric (with-me-ness) that relies on the focus of attention of the
children to assess to which extent the child is into the task.

\item \emph{The paper would benefit from further explanation of the adult's
active role throughout the robot-child interaction} The new version of the manuscript hopefully
clarifies the role of the facilitator in the different studies.

### 2016-01-04 main.tex
Rephrasing of the conclusion of the section 3.2

We find that the feedback of five out of the eight children does significantly
correlate with the actual performance of the robot (Table~\ref{table:scores}).
This indicates that these children are effectively taking on the teacher's
role and are seriously providing feedback to the robot. The observation of
the three remaining children reveal a variety of behaviors (for instance, one
was actually rating how ``nice'' the robot is, and another one had rather
``artistic'' writing style preferences that were independent of the actual
legibility of the robot), and during this specific experiment, we eventually
found that they were all seriously engaged into teaching.

To summarize, this experiment shows that children with actual handwriting
impairments do accept well the robot, can commit themselves into long
interactions, do take on the role of the teacher (as evidenced by both the
number of writing demonstrations they provided and their self-inclination to
give feedback to the robot), and finally take their teacher's role seriously.

>>>>

We find that the feedback of five out of the eight children does significantly
correlate with the actual performance of the robot (Table~\ref{table:scores}).
This indicates that these children are effectively taking on the teacher's
role and are seriously providing feedback to the robot. The observation of
the three remaining children reveal a variety of behaviors (for instance, one
was actually rating how ``nice'' the robot is, and another one had rather
``artistic'' writing style preferences that were independent of the actual
legibility of the robot), but none of these children adopted a playful-only
behavior toward the robot.

To summarize, this experiment shows that children with actual handwriting
impairments do accept well the robot, can commit themselves into long
interactions, do take on the role of the teacher (as evidenced by both the
number of writing demonstrations they provided and their self-inclination to
give feedback to the robot), and finally do usually take their teacher's role seriously.
